---
title: "Take Home 4"
author: "Serena Kongara"
date: "November 21, 2019"
output: html_document
---

```{r setup, include = FALSE}

#-----------------------------------------------------------
#
# Take Home Assignment 4
# Serena Kongara November 21st 2019 skongara@tulane.edu
#
# This script provides the code and answers to take home
# assignment 4
#
# It requires the use of the ggplot2, tidyverse, devtools, funModeling, Hmisc, and dplyr packages.
# 
# It was developed in R 3.5.1. rmarkdown is automatically handled by rstudio.
#---------------------------------------------------------

#loading required libraries
library(ggplot2)
library(tidyverse)
library(devtools)
library(funModeling)
library(Hmisc)
library(dplyr)

#set any other options---
# none
```


```{r docinfo, include = FALSE}

# print and get runtime ----
date()
# session info ----
devtools::session_info()
sessionInfo()
```

```{r data, include = FALSE}
#code for number 1 and 2
my_data <- read_csv("C:/Users/skongara/Downloads/ghana_bednet.csv")

```
1.There are 26342 observations.
2. There are 8 variables. 
```{r question 3}
#code for number 3
my_data$sex[my_data$sex == "."] <- NA
my_data$sex[my_data$sex == "F"] <- NA
describe(my_data$sex)
```
3. The proportion of male children represented in the data set is .505. 


```{r question 4}
#code for number 4

prop.table(table(my_data$sex))
follyr_complete <- my_data[complete.cases(my_data), ]
sum(follyr_complete[follyr_complete$'follyr'& follyr_complete$sex == "1", 3])

```
4. The total number of years of follow up experienced by female children in intervention areas is 328680.9. 

```{r question 5}
#code for number 5
prop.table(table(my_data$outcome == "1" & my_data$cluster ==52))
```
5. The proportion of children in cluster 52 who died during their follow up is .0003416597. 

``` {r question 6}
#code for number 6
sd(my_data$agemn)
summary(my_data$agemn)
```
6. The mean of subject's age is 25.287 months and the standard deviation is 18.0247 months. 

```{r question 7}
#code for number 7
newdata <- my_data [ which(my_data$bednet=='0'), ]
newdata <- mutate(newdata, agefollow = cut(follyr, breaks= c(0, 1, 2)))
mytable <- table(newdata$agefollow, newdata$outcome)
prop.table(mytable)
newdata1 <- my_data [which(my_data$bednet=='1'), ]
newdata1 <- mutate(newdata1, agefollow = cut(follyr, breaks= c(0, 1, 2)))
mytable1 <- table(newdata1$agefollow, newdata1$outcome)
prop.table(mytable1)
```
7.The moratlity rate in the first year of follow up for the control group is 2.8%  and in the second year is .6% .
The mortality rate in the first year of follow up for the control group is 2.4% and in the second year is .5% .


```{r question 8}
#code for number 8 
counts <- table(my_data$outcome,  my_data$cluster)
barplot (counts, main= "8. Cluster vs Outcome", xlab="Outcome", col=c("darkblue", "red") , legend = rownames(counts))
```

```{r question 9}
#code for number 9 
model <- glm(outcome ~ bednet, data = my_data, family =binomial(link="logit"))
summary(model)       
```

```{r question 10}
#code for number 10
atable <- table(my_data$bednet, my_data$outcome)
prop.table(atable)
fisher.test(table(my_data$bednet, my_data$outcome))

```
10. Being in an intervention cluster was associated with a lower probabaility of dying during the course of the trial. The odds ratio was .8358001 and the p value was .01016.


