---
title: "Exam 3"
author: "Serena Kongara"
date: "12/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#-------------------------------------------------------------------------------
#
#Exam 3
#Serena Kongara December 15th 2019 skongara@tulane.edu
#
#This script provides the code and answers to exam 3
#
#It requires the use of the ggplot2, tidyverse, devtools, funModeling, Hmisc, plyr, dplyr, datasets and car packages.
# It was developed in R 3.5.1. rmarkdown is automatically handled by rstudio.
#-------------------------------------------------------------------------------
#loading required libraries
library(ggplot2)
library(tidyverse)
library(devtools)
library(funModeling)
library(Hmisc)
library(plyr)
library(dplyr)
library(datasets)
library(car)
#set any other options---
# none
```

```{r docinfo, include = FALSE}
# print and get runtime ----
date()
# session info ----
devtools::session_info()
sessionInfo()
```

```{r data, include = FALSE}
#code for number 1 and 2
my_data <- read_csv("/Users/serenakongara/Desktop/exam3/nola_crime_2018.csv")
```

1. There are 9970 observations in the data set. 
2. There are 24 variables in the data set. 

```{r, question 3}

unique <- unique(my_data$Item_Number)
 length(unique(my_data$Item_Number))


```

3. There are 9384 unique item numbers in the data set. 

``` {r, question 4}
data2<-subset(my_data, !duplicated(my_data$Item_Number))
summary(data2)
table(data2$Signal_Description, data2$GNOCDC_LAB=="IRISH CHANNEL")
```
4. There is one aggravated crime in the Irish Channel neighborhood. 
``` {r, question 5}

question5<- count(data2, "GNOCDC_LAB")
question5 %>% arrange(desc(freq))

```
5. The two neighborhoods with the highest number of crime reports are 
the French Quarter and Little Woods. 

```{r, question 6}
crimepercity<-table(data2$GNOCDC_LAB)
print(crimepercity)
pop2<- table(data2$pop)
print(pop2)
crimeperpop<- (crimepercity*1000/pop2)    #per 1000 people
print(crimeperpop)
#you produced the no. crimes and population figures correctly but your crimepercity object is not behaving correctly to produce the correct crime rates per 1000 persons

```

```{r, question 7}
question7<- count(data2, "GNOCDC_LAB", "pop") 
question7 %>% arrange(desc(freq))
littlewoods<- subset(data2, GNOCDC_LAB == 'LITTLE WOODS')
describe(littlewoods)

```
7. The largest neighborhood by population in the dataset is Little Woods. Its population is 44947.

```{r, question 8}
#to find the number of unique crimes
summary(data2$GNOCDC_LAB== "FLORIDA DEV")
#to find the population size
floridadev <- subset(data2, GNOCDC_LAB == 'FLORIDA DEV')
describe(floridadev)

#lowest crime rates was Lakewood
```
8. According to the table from question 6, Florida Dev has the lowest crime rate. The population size of Florida Dev is  4976 and the number of unique crimes is 6. 

```{r, question 9}
LAKEWOOD <- count(data2$GNOCDC_LAB == "LAKEWOOD")
print(LAKEWOOD)
#10 unique crimes in lakewood
question9<- table(data2$GNOCDC_LAB)/10
print(question9)
#since we are wanting to produce rate ratios, the number of unique crimes in lakewood wouldn't be the only measure to include in the denominator as part of the reference ratio (as you indicate above in your code); you would produce a rate ratio for each neighborhood, therefore, you would include in the numerator the crime rate for neighborhood 1, etc. and crime rate for Lakewood in the denominator

```



```{r, question 10}
aggravated<- data2 %>% filter(GNOCDC_LAB == "FRENCH QUARTER") %>%
print(aggravated)
#15 aggravated crimes in the French Quarter


#data only includes aggravated crimes
newdata <- subset(data2, Signal_Type %in% c('60D','60','55','51','42', '42U', '42M', '37D','37','34S', '34D', '34C','34')) 
question10 <-table(newdata$GNOCDC_LAB)/15
print(question10)

#number of aggravated crimes in the French Quarter is correct, however, you rate ratios are still not correct due to the same issue in the previous question of how to set-up the calculation for a ratio ratio


```

```{r, question 11}
 
question11<-count(data2$Signal_Description)
question11 %>% arrange(desc(freq))

question11b <- count(data2$Charge_Description)
question11b %>% arrange(desc(freq))
```
11. The most common cause for a police report to be filed is Disturbance(Domestic) and the most common charge to be filed is Theft. 

``` {r, question 12}

str(data2$pop)

question12 <-table(data2$Signal_Description== "DISTURBANCE (DOMESTIC)", data2$GNOCDC_LAB)
print(question12)
#this yielded the correct number of domestic disturbances by neighborhood; however we are interested in the rate (not proportion)

prop.table(question12)*100 #to make percentages

```


```{r, question 13}

agepercity<-table(data2$Victim_Age,data2$GNOCDC_LAB)
print(agepercity)
simple.fit <- lm(crimepercity~agepercity,  data=data2)
summary(simple.fit)
#error found: variable lengths differ for variable agepercity, and as a result the object simple.fit was not found; model didn't run

```


```{r, question 14}
question14 <-count(data2, 'GNOCDC_LAB')
a<- question14 %>% arrange(desc(freq))
question14a<- count(data2,'GNOCDC_LAB', 'pop')
b<-question14a %>% arrange(desc(freq))
full_join(a,b)

#your numerator is correct, but your denominator (population figures) are extremely high; the correct answer is Lakewood as the neighborhood likely to have the worst police responsiveness (10 crime incidents, population= 13989, crime rate=0.715)

```
14. The location with the highest pop and lowest reported crime is VILLAGE DE LEST. The population is	1882930 and there are only 74 crimes.

``` {r, question 15}
poisson <- rpois(10000, lambda=4.7)
hist(poisson, main="A histogram of a Poisson distribution")

```

```{r, question 16}
data(iris)
summary(iris)

answer <- list()
for(n in names(split)){   
   data <- split[[n]]   
   answer[[n]] <- data.frame(species=n,
                          mean.sepal.length=mean(data$Sepal.Length))
}
print(answer)
```

16. The mean sepal length for the setosa is 5.006, for the versicolor 5.936, and for the virginica 6.588.

``` {r, question 17}
data(mtcars)
summary(mtcars)

ggplot(mtcars, aes(x=mpg, y=disp, color=cyl)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)

```


``` {r, question 18}
vector18 <- c(-4.89, -1.93, -1.11, 3.94, 0.46, -3.85, -0.20, 6.04,9.36, 7.26, 4.88, 13.45, -2.93, 6.39, -16.22, -3.24, 6.86, 11.87, 1.81, -2.54)


zscore <- function(x) {
   
   zscore = (x - mean(x))/sd(x)
  return(zscore)}
  
zscore(vector18) 

```
